\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2016-05-23 mSystems/mSys_intro.tex
two more cites in introduction, minor changes in structure

The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research. We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism of the bile acids \cite{bileacids}, of the choline \cite{choline} or key-route metabolites as short-chain fatty acids \cite{scfa1, scfa2} which is also involved in immune system maturation \cite{scfa3, scfa4}. Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections as obesity \cite{ob1, ob2}, malnutrition \cite{nutr} among other multiple diseases \cite{Moya_trends}. High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body. Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health}. To define normal microbiota and how it's compositional changes can origin some diseases are important issues still in need for scientific answers \cite{normal1, normal2}.

@@

Here we present the imprints of health status (healthy or disease) in macroscopic properties of microbiota, by studying the temporal variability. We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic therapy and healthy individuals. Having seen that all cases follows Taylor's law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as it was applied recently by Blumm {\it et al.} \cite{ranking}. We use this mathematical framework to explore the temporal stability of the microbiota in different conditions in order to understand how this affects the healthy status of the subjects. Finally, we have engineered a complete software framework, {\it ComplexCruncher}, to carry out the analysis of the temporal dynamics of microbiotas, which is ready to be implemented by other users.

>>>>

The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research. As evidence of new concepts 'holobiont' and 'hologenome' is increasing each day \cite{holo1, holo2}, research not only focus on the human physiology but also on the microbial population that surround ourselves. We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism of the bile acids \cite{bileacids}, of the choline \cite{choline} or key-route metabolites as short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}. Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections as obesity \cite{ob1, ob2}, malnutrition \cite{nutr} among other multiple diseases \cite{Moya_trends}. High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body. Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health}. To define normal microbiota and how it's compositional changes can origin some diseases are important issues still in need for scientific answers \cite{normal1, normal2}.

@@

Here we present the imprints of health status (healthy or disease) in macroscopic properties of microbiota, by studying its temporal variability. We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic therapy and healthy individuals. Having seen that all cases follows Taylor's law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as it was applied recently by Blumm {\it et al.} \cite{ranking}. We use this mathematical framework to explore the temporal stability of the microbiota in different conditions in order to understand how this affects the healthy status of the subjects. Finally, we have engineered a complete software framework, {\it ComplexCruncher}, to carry out the analysis of the temporal dynamics of microbiotas, which is ready to be implemented by other users.
